---
title: "COMPSCIX 415.2 Homework 9/Final"
author: "Chris Boyd"
date: "3/31/2019"
output:
  html_document:
    self_contained: true
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
```

# Exercise 1

> Sampling Distributions, Functions and For Loops (10 points)

> Function to return calculated mean and standard deviation

```{r}

sample_fun <- function(samp_size, samp_rate) {
  
  data <- rexp(n = samp_size, rate = samp_rate)
  samp_avg = mean(data)
  samp_std_dev = sd(data) 
  
  stats <- list(samp_avg = samp_avg, samp_std_dev = samp_std_dev)
  return(stats)
}

```

> Sample size = 50

```{r}

n <- 1000

samp_size = 50
samp_rate = 1/10000


sample_means <- rep(NA, n)
sample_sds <- rep(NA, n)

for(i in 1:n) {
    
  stats <- sample_fun(samp_size, samp_rate)
  sample_means[i] <- stats$samp_avg 
  sample_sds[i] <- stats$samp_std_dev
    
}

samples <- tibble(sample_means, sample_sds)

ggplot(samples) +
  geom_histogram(aes(x = sample_means))

sd_of_means <- sd(samples$sample_means)

theoretical_se <- 1/samp_rate / sqrt(samp_size)

mean_of_sds <- mean(samples$sample_sds)

emperical_sd <- mean_of_sds / sqrt(samp_size)

sd_of_means
theoretical_se
emperical_sd

```

> Sample size = 5000

```{r}

n <- 1000

samp_size = 5000
samp_rate = 1/10000


sample_means <- rep(NA, n)
sample_sds <- rep(NA, n)

for(i in 1:n) {
    
  stats <- sample_fun(samp_size, samp_rate)
  sample_means[i] <- stats$samp_avg 
  sample_sds[i] <- stats$samp_std_dev
    
}

samples <- tibble(sample_means, sample_sds)

ggplot(samples) +
  geom_histogram(aes(x = sample_means))

sd_of_means <- sd(samples$sample_means)

theoretical_se <- 1/samp_rate / sqrt(samp_size)

mean_of_sds <- mean(samples$sample_sds)

emperical_sd <- mean_of_sds / sqrt(samp_size)

sd_of_means
theoretical_se
emperical_sd

```

# Exercise 2 - Linear Regression (5 points)

> Use the broom package to output the coefficients and the R-squared

```{r}
import_file_path = "../data/train.csv"
train <- read.csv(import_file_path, header = TRUE)

mod_1 <- lm(SalePrice ~ LotArea + OverallQual + ExterQual, data = train)

glance(mod_1)
tidy(mod_1)

ggplot(train) +
  geom_point(mapping = aes(x=SalePrice, y=LotArea))

```

`

> Interpret the coefficient on LotArea

For every extra square foot in the LotArea, the sale price increases, on average, by $1.45, with all other features being held constant.  However, there are some pretty major outliers.  If you filter by LotArea < 45000, you get an increase of $4.10 with each increase in sq ft of lotsize.  That data is below.

> Interpret the coefficient on ExterQualGd

The mean difference in the sale price between ExterQualGd and ExterQualEx is -$71k (or -$70k with the LotArea filter.)


> Compare this model to the model we fit in HW 7 with GrLivArea, OverallQual, Neighborhood. Which is the better fitting model?

Our HW7 model had an adjusted R-squared of .78, which was a good fit.  This model has an adjusted R-squared of .69 (or .72 with the LotArea filter), which is less of a good fit.   HW7 is a better fitting model. 

# Exercise 3 - AB Testing (5 points)


```{r}

### filter out the outliers

train_filtered <- filter(train, LotArea < 45000)

mod_filtered <- lm(SalePrice ~ LotArea + OverallQual + ExterQual, data = train)

glance(mod_filtered)
tidy(mod_filtered)

```



```{r}
import_file_path = "../data/ab_test_data.csv"
ab_test_data <- read.csv(import_file_path, header = TRUE)
```

> What proportion of visitors converted for each version of the webpage?

```{r}
a_total <- filter(ab_test_data, version == 'A')
a_converted <- filter(a_total, conversion == 1)

b_total <- filter(ab_test_data, version == 'B')
b_converted <- filter(b_total, conversion == 1)

a_conversion <- count(a_converted) / count(a_total)
b_conversion <- count(b_converted) / count(b_total)

count(a_total)
a_conversion
count(b_total)
b_conversion

```

```{r}

true_a <- .0415
true_b <- .1
n_a <- 2000
n_b <- 2000

samp_a <- rbinom(n = 1, size = n_a, prob = true_a)
samp_b <- rbinom(n = 1, size = n_b, prob = true_b)

two_prop_test <- prop.test(c(samp_a, samp_b), c(2000, 2000))

two_prop_test$p.value


```

The p-value tells us that the conversion rates for Version A and B are significantly different than each other, so we'd conclude here that whatever changes were made to the did have an effect.