---
title: "COMPSCIX 415.2 Homework 5"
author: "Chris Boyd"
date: "2/26/2019"
output:
  html_document:
    self_contained: true
---

```{r message=FALSE, warning=FALSE}
library(mdsr)
library(dplyr)
library(tidyverse)
```

# Table of Contents

1. Tidyverse packages
2. R Basics
3. Data Import / Export
4. Visualization
5. Data munging and wrangling
6. EDA
7. Git and Github

# Section 1: The tidyverse packages

> Can you name which package is associated with each task below?

* Plotting: **ggplot2**
* Data munging/wrangling: **dplyr**
* Reshaping (speading and gathering) data: **tidry**
* Importing/exporting data: **readr**

> Now can you name two functions that you’ve used from each package that you listed above for these tasks?

* Plotting - ggplot2: **geom_point()** and **aes()**
* Data munging/wrangling: dplyr: **filter()** and **summarise()**
* Reshaping (speading and gathering) data - tidry: **gather()** and **spread()**
* Importing/exporting data - readr: **read_csv()** and **write_delim()**

# Section 2: R Basics

> Fix this code with the fewest number of changes possible so it works:

> 1. My_data.name___is.too00ooLong! <- c( 1 , 2   , 3 )

```{r}
# corrected with fewest possible changes:
My_data.name___is.too00ooLong <- c( 1 , 2   , 3 )
My_data.name___is.too00ooLong

# corrected with a few more changes:
my_data <- c( 1 , 2   , 3 )
my_data

```

> 2. my_string <- C('has', 'an', 'error', 'in', 'it)

```{r}
# corrected with fewest possible changes:
my_string <- c('has', 'an', 'error', 'in', 'it')

```

> 3. Look at the code below and comment on what happened to the values in the vector.

```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector
typeof(my_vector)
```

Wrapping the 3 and 4 in the single quote cast each element in the vector to a 'character' datatype.


# Section 3: Data import/export 

> 1. Download the rail_trail.txt file from Canvas (in the Midterm Exam section) and successfully import it into R. Prove that it was imported successfully by including your import code and taking a glimpse of the result.

```{r}
import_file_path = "../data/rail_trail.txt"
rail_trail <- read.csv(import_file_path, sep ="|", header = TRUE)
glimpse(rail_trail)
```

> 2. Export the file into a comma-separated file and name it “rail_trail.csv”. Make sure you define the path correctly so that you know where it gets saved. Then reload the file. Include your export and import code and take another glimpse.

```{r}
export_file_path = "../data/rail_trail.csv"

write.csv(rail_trail, file = export_file_path, row.names=FALSE)

import_file_path_csv = "../data/rail_trail.csv"
rail_trail_csv <- read_csv(import_file_path_csv, col_names = TRUE)
glimpse(rail_trail_csv)

```

# Section 4: Visualization

> Critique this graphic: give only three examples of what is wrong with this graphic. Be concise.

* The segments overlap.   Each 'Men/Women' segment is composed of each 'Age' segemnt.   A grouping function would help.
* The populations sizes aren't indicated.   Would be nice to know how big each group is.
* Color scheme is weird.  Age is all black, while gender uses a color gradient.



> Reproduce this graphic using the diamonds data set.

```{r}
ggplot(data = diamonds) +
  geom_boxplot(position="identity", aes(x = cut, y=carat, fill=color)) +
  coord_flip() + 
  xlab("CARAT OF DIAMOND") +
  ylab("CUT OF DIAMOND")
```

The easiest way to make this more readable is to omit the 'position="identity"' argument.
```{r}
ggplot(data = diamonds) +
  geom_boxplot(aes(x = cut, y=carat, fill=color)) +
  coord_flip() + 
  xlab("CARAT OF DIAMOND") +
  ylab("CUT OF DIAMOND")
```

Another way is to add a facet_wrap.   I think the plot above is more readable, however.

```{r}
ggplot(data = diamonds) +
  geom_boxplot(aes(x = cut, y=carat, fill=color)) +
  coord_flip() + 
  facet_wrap(~ color) +
  xlab("CARAT OF DIAMOND") +
  ylab("CUT OF DIAMOND")
```

# Section 5: Data munging and wrangling

> Is this data “tidy”? If yes, leave it alone and go to the next problem. If no, make it tidy. Note: this data set is called table2 and is available in the tidyverse package. It should be ready for you to use after you’ve loaded the tidyverse package.

No.   Type and count should be spread to make it tidy data.

```{r}
table2 %>%
  spread(key = type, value = count)
```

> Create a new column in the diamonds data set called price_per_carat that shows the price of each diamond per carat (hint: divide). Only show me the code, not the output.

```{r}
new_diamonds = mutate(diamonds, price_per_carat = price / carat)
```

> For each cut of diamond in the diamonds data set, how many diamonds, and what proportion, have a price > 10000 and a carat < 1.5? There are several ways to get to an answer, but your solution must use the data wrangling verbs from the tidyverse in order to get credit.

```{r}
diamonds %>%
    group_by(cut) %>%
      summarise(
        total = n(),
        total_count = count(price > 10000 & carat < 1.5),
        proportion  = total_count / total 
      )
      

```

# EDA