---
title: "Homework 6"
author: "Chris Boyd"
date: "3/9/2019"
output:
  html_document:
    self_contained: true
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(mosaicData)
```

# Exercise 1

> What variables are in this data set?

The variables are: outcome, smoker, age

> How many observations are there and what does each represent?

There are 1,314 observations.  Each observation represents a woman who has been classified as a smoker or non-smoker, with a status of dead or alive 20 years after the original survey.

> Create a table (use the R code below as a guide) and a visualization of the relationship between smoking status and outcome, ignoring age. What do you see? Does it make sense?

```{r}
Whickham %>% 
  count(outcome, smoker) 
```
```{r}
Whickham %>% 
  ggplot() +
    geom_bar(mapping = aes(x=outcome, fill=smoker))

```

The data doesnt make sense.   It looks like there are more dead non-smokers than dead smokers.   I would expect the inverse.   Same goes for alive smokers/non-smokers.

> Recode the age variable into an ordered factor with three categories: age <= 44, age > 44 & age <= 64, and age > 64. Now, recreate visualization from above, but facet on your new age factor. What do you see? Does it make sense?

```{r}
Whickham <- Whickham %>% 
  mutate(age_fct = case_when(age >= 0 & age < 40 ~ '< 40',
                               age >= 40 & age < 60 ~ 'between 40 and 60',
                               age >= 60 ~ '> 60'))
Whickham <- Whickham %>% mutate(age_fct = factor(age_fct, ordered=TRUE))
Whickham <- Whickham %>% mutate(age_fct = fct_relevel(age_fct, '< 40', 'between 40 and 60', '>60'))

Whickham %>% 
  ggplot() +
    geom_bar(mapping = aes(x=outcome, fill=smoker)) +
    facet_wrap(~ age_fct)

```

This makes more sense.   Here I see that the older you get, the more likely you are to be dead if you were a smoker.    There is a slight difference (favoring dead) between dead/alive in the largest "<40" segment, which has a huge impact on the overall percentage of "deads."

# Exercise 2

>The Central Limit Theorem states that the sampling distribution of sample means is approximately Normal, regardless of the distribution of your population. For this exercise our population distribution will be a Gamma(1,2) distribution, and we’ll show that the sampling distribution of the mean is in fact normally distributed.

>Generate a random sample of size n = 10000 from a gamma(1,2) distribution and plot a histogram or density curve. Use the code below to help you get your sample.

```{r}
n <- 10000

gamma_samp <- tibble(x = rgamma(n, shape = 1, scale = 2)) 

gamma_samp %>%
  summarise(
    mean(x, na.rm=TRUE),
    sd(x, na.rm=TRUE)
  )

 

```

> Pretend the distribution of our population of data looks like the plot above. Now take a sample of size n = 30 from a Gamma(1,2) distribution, plot the histogram or density curve, and calculate the mean and standard deviation.

```{r}

sample_size <- 30
gamma_samp <- tibble(x = rgamma(sample_size, shape = 1, scale = 2))


gamma_samp %>% 
  ggplot() +
    geom_histogram(aes(x = x)) +
    geom_vline(aes(xintercept=mean(x)),
            color="blue", linetype="dashed", size=1) + 
    geom_vline(aes(xintercept=sd(x)),
            color="red", linetype="dashed", size=1)  

  
```

> Take a sample of size n = 30, again from the Gamma(1,2) distribution, calculate the mean, and assign it to a vector named mean_samp. Repeat this 10000 times!!!! The code below might help.

```{r}

sample_size <- 30
mean_samp <- rep(NA, 10000)

# start a loop
for(i in 1:10000) {
  g_samp <- rgamma(sample_size, shape = 1, scale = 2)
  mean_samp[i] <- mean(g_samp)
}

# Convert vector to a tibble
mean_samp <- tibble(mean_samp)

mean_samp %>%
  summarise(
    mean=mean(mean_samp),
    sd=sd(mean_samp)
  )

mean_samp %>% 
  ggplot() +
    geom_histogram(aes(x = mean_samp)) +
    geom_vline(aes(xintercept=mean(mean_samp)),
            color="blue", linetype="dashed", size=1) + 
    geom_vline(aes(xintercept=sd(mean_samp)),
            color="red", linetype="dashed", size=1)  


  
```
> Did anything surprise you about your answers to #6?

I'm suprised the SD is so small.


```{r}

sample_size <- 300
mean_samp <- rep(NA, 10000)

# start a loop
for(i in 1:10000) {
  g_samp <- rgamma(sample_size, shape = 1, scale = 2)
  mean_samp[i] <- mean(g_samp)
}

# Convert vector to a tibble
mean_samp <- tibble(mean_samp)

mean_samp %>%
  summarise(
    mean=mean(mean_samp),
    sd=sd(mean_samp)
  )

mean_samp %>% 
  ggplot() +
    geom_histogram(aes(x = mean_samp)) +
    geom_vline(aes(xintercept=mean(mean_samp)),
            color="blue", linetype="dashed", size=1) + 
    geom_vline(aes(xintercept=sd(mean_samp)),
            color="red", linetype="dashed", size=1)  

  
```
No, increasing the sample size to 300 has no major change on the SD. 


